Competing Constructions for German Measure Phrases

### 2017-03-29 Latex/article.tex
corpus finished; exp 1 progress

Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
%First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
%
%\begin{exe}
%  \ex\label{ex:analyticalpartitive} 
%  \begin{xlist}
%    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
%                                                a glass of the red wine\\
%                                                \glt a glass of the red wine
%       }
%    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
%                                                a glass of red wine\\
%                                                \glt a glass of red wine
%       }
%  \end{xlist}
%\end{exe}
%
If the measure noun denotes a container or vessel, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).

@@

These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
There are no clear hypotheses about such interactions, however.
Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container or vessel measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.

@@

This section briefly reviews existing analyses of the \PGCa-\NACa\ alternation and related issues.
I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
%While the constructions and the alternation are standardly covered in descriptive grammars such as \citet[XYZ]{Eisenberg2013a}, the research literature contains little specific discussion of the situation in German.
I discuss four main factors that might influence the alternation:
first, the different syntactic structures of the variants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
third, a semantic prototypicality effect of the degrees of grammaticalisation (plurality);
fourth, preferences in different registers.

@@

In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
It contains just over one billion tokens.}
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
%\footnote{They behave like mass nouns mostly in that they occur without a determiner in the singular.}
The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.

@@

In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
%The scripts generating the actual queries were crafted so as to make sure that adjectives and nouns were searched in the appropriate forms.
%The CQL query in (\ref{ex:data:query01}) shows an example query for a neuter mass noun like \textit{Wasser} `water' which finds relevant sequences such as the ones shown in (\ref{ex:data:query02}).%
%\footnote{The COW creators used to make corpora available in the IMS Open Corpus Query Workbench \citep{EvertHardie2011}, which has a query syntax informally referred to as CQL.
%Documentation can be obtained at \url{http://cwb.sourceforge.net}.}
%The versions with the genitive singular \textit{Liters} and the dative plural \textit{Litern} are not shown in (2) for space reasons but are also found by the query in (1).
%In terms of context, single sentences containing the target were exported.
%
%\begin{exe}
%  \ex{\label{ex:data:query01} \texttt{\small[ lemma="Liter" ] [ pos="ADJA" \& word=".+(en|em|es)" ] [ lemma="Wasser" ]}}
%  \ex\label{ex:data:query02}
%  \begin{xlist}
%    \ex{\label{ex:data:query02a} Liter [klaren Wassers]\Sub{Gen}}
%    \ex{\label{ex:data:query02b} Liter [klares Wasser]\Sub{Nom\slash Acc}}
%    \ex{\label{ex:data:query02c} Liter [klarem Wasser]\Sub{Dat}}
%  \end{xlist}
%\end{exe}
In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 1,876 sentences.
%Bootstrapping measure nouns first was more efficient in the plural for reasons to complicated to discuss here, especially because the data was not properly analysed anyway.
As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns, and this sample was not used for full analysis.
%Consequently, it does not make sense to model the influencing factors in the same way as for singular kind nouns.
}

@@

In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two variants.
To capture this, class information was annotated for measure lemmas.
The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the frequencies of the respective nouns, very coarse classes were necessary.
With typical examples and the number of data points in the final sample, the classes are:
\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; size 1,968),
\textit{Container} (\textit{Eimer} `bucket'; size 740), 
\textit{Amount} (\textit{Menge} `amount'; size 1,364), 
\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; size 713).
The few lemmas that did not fit into either of these classes were labeled \textit{Rest} (size 278).

@@

In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation.
The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
The regressors listed in Table~\ref{tab:variables} with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
The same goes for those listed with the kind lemma as their unit of reference.
The results of the estimation are shown in Table~\ref{tab:bigtable}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

\begin{sidewaystable}
  \centering
    \begin{tabular}{lllrrrrrrrcc}
    Level     & Regressor         & Level     & \pPB  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
    &                   &           &       & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule

    First     & Badness           & (numeric) & 0.001 & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
              & Genitives         & (numeric) & 0.001 & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
              & Leftcontext       & D         & 0.001 &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
              &                   & O         &       &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
              &                   & A         &       &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
              &                   & P         &       &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
              & Measurecase       & Acc       & 0.001 &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
              &                   & Dat       &       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
	      & Measurenumber     & Pl        & 0.001 &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]

    Second    & Kindattraction    & (numeric) & 0.033 &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
    (Kind)    & Kindfreq          & (numeric) & 0.066 &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
              & Kindgender        & Neut      & 0.001 & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
	      &                   & Fem       &       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]

    Second    & Measureattraction & (numeric) & 0.015 &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
    (Measure) & Measureclass      & Rest      & 0.001 &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
              &                   & Container &       &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
              &                   & Amount    &       &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
              &                   & Portion   &       &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
              & Measurefreq       & (numeric) & 0.003 & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
  \end{tabular}
  \smallskip
  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); for details see text}
  \label{tab:bigtable}
\end{sidewaystable}

Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
Generalized variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.423$ and $R^2_c=0.515$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
The rate of correct predictions is $0.855$, which means a proportional reduction of error of $\lambda=0.354$.
The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$

The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (for the upper and lower 2.5th percentiles).
The column \textit{CI contains 0} shows an asterisk for those intervals that do not include 0 as a more modest replacement of significance levels.
Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
All regressors have low \pPB\ values of magnitude $10^{-2}$ and $10^{-3}$, and only \textit{Kindattraction} ($p_{\text{PB}}=0.033$) and \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').

Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
Both the MN model and the F model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
The algorithm was run using 4 chains and 1,000 iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
\footnote{Consider that with a coefficient of 10, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22026.47$.
To reliably estimate such coefficients, extremely large samples would be required.}
The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
Since both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}), I only interpret the MLE model in the next section.

@@

The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
Before I turn to some details, it should be noticed that the results from standard MLE (with bootstrap) and MCMC (with default priors) obviously differ in numerical details -- but not enough to result in different interpretations.
The signs and magnitudes of the coefficients are identical with the two methods, and the confidence intervals have roughly the same width and symmetry properties.
In only one single case (MN model, \textit{Measureattraction}) the MLE bootstrap confidence interval does not include $0$ while the MCMC interval does (barely, with $-0.01$).
While I only interpret the MLE results, virtually the same interpretation could be made for the MCMC results.
Furthermore, the MN model and the F model agree in many relevant aspects.
Disagreement between the models will be discussed below.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.75\textwidth]{../R/output/corpus_Measureattraction}
  \caption{Effect plot for the regressor \textit{Measureattraction}}
  \label{fig:eff:measureattraction}
\end{figure}

First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:measureattraction}.%
\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis),assuming typical values of all other regressors.}
The effect is mostly as expected:
if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
Only the F model \textit{Kindattraction} regressor is clearly useless (`non-significant') in the PBmodcomp test, and the effect plot shows a corresponding picture.
This is related to the fact that in the F model the kind lemma effect and the corresponding second-level model could not be estimated successfully (remember that $\sigma_{\text{Kindlemma}}=0.000$).
The effect plots show a picture in line with a \textit{divide by 4} interpretation of the coefficients as shown in Table~\ref{tab:bigtable}.%
\footnote{Coefficients are on the logit scale, which does not linearly correspond to the probability scale.
If coefficients are reasonably small, however, dividing the coefficient by $4$ gives a good approximation of the change on the probability scale for each increment of the response variable \citep[82]{GelmanHill2006}.}
Given that all regressors have been z-transformed, we can make statements such as the following:
If a masculine kind noun is one standard deviation more frequently encountered in the \PGCd\ (compared to its frequency in the \NACb) than the kind nouns on average, the probability that the \PGCa\ is chosen instead of the \NACa\ with this kind noun in it rises by $\beta^{MN}_{\text{Kindattraction}}/4=0.49/4=0.1225$, etc.
For measure nouns in MNPs with masculine kind nouns, it is $\beta^{MN}_{\text{Measureattraction}}/4=0.21/4=0.0525$, and for measure nouns in MNPs with feminine kind nouns, it is $\beta^{F}_{\text{Measureattraction}}/4=0.43/4=0,1075$.
As readers can observe, these numbers correspond approximately to the slopes of the effect plots.

@@

I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
Plurality has a clear effect in the expected direction (plural favours the \PGCa) in the MN model ($p=0.002$ for \textit{Measurenumber}).
In the F model, the effect has the same direction, but the estimate is quite imprecise ($p=0.160$), again likely due to the low sample size.
Figure~\ref{fig:eff:leftcontext} shows, however, that the left context of the measure noun influences the choice of the variant (in both models, the regressor has $p=0.002$).
Cardinals have a strong tendency to co-occur with the \NACa\ in both models.

>>>>

Finally, (\ref{it:intro:other}) suggests that there are alternative ways of expressing similar quantificational meanings.
%First of all, the analytic (pseudo-)partitive with \textit{von} `of' is only available as an alternative to the PGC if the kind noun phrase contains a (definite or indefinite) determiner as in (\ref{ex:analyticalpartitive}).
%
%\begin{exe}
%  \ex\label{ex:analyticalpartitive} 
%  \begin{xlist}
%    \ex[ ]{\label{ex:analyticalpartitive:a}\gll ein Glas von dem roten Wein\\
%                                                a glass of the red wine\\
%                                                \glt a glass of the red wine
%       }
%    \ex[*]{\label{ex:analyticalpartitive:b}\gll ein Glas von rotem Wein\\
%                                                a glass of red wine\\
%                                                \glt a glass of red wine
%       }
%  \end{xlist}
%\end{exe}
%
If the measure noun denotes a container, constructions with \textit{voll} (\textit{von})\slash\textit{voller} `full (of)' and \textit{mit} `with' are available, see (\ref{ex:alternatives}).

@@

These construction have very idiosyncratic properties (the construction with \textit{voller} is discussed in \citealp{Zeldes2018}, the construction with \textit{mit} is discussed in \citealp{Bhatt1990}), they are not semantically equivalent to the NAC and the PGC, and -- most importantly -- they are only available for a small subset of measure nouns.
Subtle interactions between these constructions and NAC\slash PGC cannot be excluded, especially because the former might act as a (partial) alternative and thus reduce the count of exemplars of the NAC, the PGC, or both for container measure nouns.
There are no clear hypotheses about such interactions, however.
Nevertheless, the corpus-based models in Section~\ref{sec:corpusstudies} are specified in a way that they could detect preferences for either NAC or PGC for the class of container measure nouns, for which the alternative constructions illustrated in (\ref{ex:alternatives}) are available.

@@

This section briefly reviews existing analyses of the \PGCa-\NACa\ alternation and related issues.
I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:externalvalidation}.
I discuss four main factors that might influence the alternation:
first, the different syntactic structures of the variants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
third, a semantic prototypicality effect of the degrees of grammaticalisation (plurality);
fourth, preferences in different registers.

@@

In step (\ref{enum:data:step1}), I exported a list of all nouns in the DECOW14A01 sub-corpus sorted by their token frequency and manually went through it from the most frequent noun downwards, selecting the first one hundred mass nouns that occurred in the list.%
\footnote{DECOW14A01 is the first slice (roughly a twentieth) of the complete DECOW14A corpus.
It contains just over one billion tokens.}
Abstract nouns which partially behave like mass nouns (like \textit{Spaß} `fun’ or \textit{Gefahr} `danger’) were excluded because they are usually not quantified in the same way as concrete mass nouns.
The hundredth selected mass noun was \textit{Schmuck} `jewelry’, which is the 3,054th most frequent noun in the original frequency list.

@@

In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for mass nouns occurring less than one hundred times, but randomly sampled one hundred hits for each mass noun that occurred one hundred or more times.
The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
Given the careful bootstrapping and sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were bootstrapped and queried, resulting in a sample of 1,876 sentences.
As stated in Section~\ref{sec:germanmeasurenps}, the NAC is virtually never used with plural kind nouns, and this sample was not used for full analysis.
}

@@

In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two variants.
To capture this, class information was annotated for measure lemmas.
The classification was designed inspired by the list in \cite[530]{Koptjevskaja2001}, but due to the low frequencies of many of the potential classes, a very coarse classification was finally used.
With typical examples and the number of data points in the final sample, the classes are:
\textit{Physical} (abstract precisely measurable units such as \textit{Liter} `litre', \textit{Meter} `metre', \textit{Gramm} `gram'; size 1,968),
\textit{Container} (\textit{Eimer} `bucket'; size 740), 
\textit{Amount} (\textit{Menge} `amount'; size 1,364), 
\textit{Portion} (natural portions like \textit{Happen} `bite' or \textit{Krümel} `crumb'; size 713).
The few lemmas that did not fit into either of these classes were labeled \textit{Rest} (size 278).

@@

\begin{figure}[hb!]
  \centering
  \includegraphics[width=0.85\textwidth]{../R/output/corpus_fixeffs}
  \caption{Coefficients (MLE) with 95\% parametric bootstrap confidence intervals; intercept (\textit{Leftcontext=C}, \textit{Measurecase=Nom}, \textit{Measurenumber=Sg}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}; 0 for all numeric z-transformed regressors) is -4.328}
  \label{fig:fixeffs}
\end{figure}

In this section, I report the results of fitting a multilevel model to the data using R \citep{R}, \textit{lme4} \citep{lme4} for Maximum Likelihood Estimation, and \textit{rstanarm} \citep{rstanarm} for Markov-Chain Monte Carlo estimation.
The purpose is to model the influence of the regressors specified in Table~\ref{tab:variables} on the probability that either the \NACa\ or the \PGCa\ is chosen.
All regressors from Table~\ref{tab:variables} were included to begin with, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was $n=5,063$ with $1,1344$ cases of \PGCa\ and $3,929$ cases of \NACa.
The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \cite[265--269,302--304]{GelmanHill2006}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

\begin{sidewaystable}
  \centering
    \begin{tabular}{llrlrrrrrrcc}
      Level & Regressor & \multicolumn{1}{l}{\pPB} & Level  & \multicolumn{2}{l}{Coefficient} & \multicolumn{2}{l}{CI Low} & \multicolumn{2}{l}{CI High} & \multicolumn{2}{l}{CI contains 0} \\
              &                   &       &          & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} & \multicolumn{1}{l}{MLE} & \multicolumn{1}{l}{MCMC} \\\midrule

    First     & Badness           & 0.001 & (numeric) & -0.165 & -0.169 & -0.264 & -0.264 & -0.074 & -0.075 &      * &      * \\
              & Genitives         & 0.001 & (numeric) & -0.630 & -0.645 & -0.710 & -0.739 & -0.523 & -0.552 &      * &      * \\
              & Leftcontext       & 0.001 & D         &  1.178 &  1.173 &  0.751 &  0.784 &  1.548 &  1.546 &      * &      * \\
              &                   &       & O         &  1.561 &  1.588 &  1.093 &  1.179 &  1.919 &  2.007 &      * &      * \\
              &                   &       & A         &  2.300 &  2.340 &  1.872 &  1.959 &  2.644 &  2.726 &      * &      * \\
              &                   &       & P         &  2.353 &  2.371 &  1.693 &  1.766 &  2.900 &  2.978 &      * &      * \\
              & Measurecase       & 0.001 & Acc       &  0.065 &  0.070 & -0.130 & -0.119 &  0.268 &  0.269 &        &        \\
              &                   &       & Dat       &  0.707 &  0.735 &  0.413 &  0.480 &  0.959 &  1.002 &      * &      * \\
	      & Measurenumber     & 0.001 & Pl        &  0.814 &  0.825 &  0.534 &  0.574 &  1.080 &  1.079 &      * &      * \\[0.5\baselineskip]

    Second    & Kindattraction    & 0.033 & (numeric) &  0.204 &  0.216 &  0.030 &  0.014 &  0.366 &  0.412 &      * &      * \\
    (Kind)    & Kindfreq          & 0.066 & (numeric) &  0.161 &  0.173 & -0.008 &  0.009 &  0.333 &  0.350 &        &      * \\
              & Kindgender        & 0.001 & Neut      & -0.007 & -0.022 & -0.407 & -0.424 &  0.408 &  0.379 &        &        \\
	      &                   &       & Fem       &  1.357 &  1.379 &  0.831 &  0.872 &  1.808 &  1.897 &      * &      * \\[0.5\baselineskip]

    Second    & Measureattraction & 0.015 & (numeric) &  0.247 &  0.275 &  0.058 &  0.078 &  0.413 &  0.476 &      * &      * \\
    (Measure) & Measureclass      & 0.001 & Rest      &  0.564 &  0.488 & -0.140 & -0.309 &  1.197 &  1.236 &        &        \\
              &                   &       & Container &  0.609 &  0.618 &  0.050 &  0.018 &  1.145 &  1.215 &      * &      * \\
              &                   &       & Amount    &  0.927 &  0.959 &  0.184 &  0.189 &  1.566 &  1.718 &      * &      * \\
              &                   &       & Portion   &  1.472 &  1.503 &  0.916 &  0.956 &  1.938 &  2.064 &      * &      * \\
              & Measurefreq       & 0.003 & (numeric) & -0.257 & -0.264 & -0.413 & -0.435 & -0.098 & -0.099 &      * &      * \\
  \end{tabular}
  \smallskip
  \caption{Coefficient table comparing Maximum Likelihood Estimation (MLE, with 95\% bootstrap confidence interval) and `Bayesian' Markov-Chain Monte Carlo estimation (MCMC); for details see text}
  \label{tab:bigtable}
\end{sidewaystable}

Standard diagnostics for the Maximum Likelihood Estimation (MLE) show that the model quality is quite high.
Generalized variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than $2$.
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.423$ and $R^2_c=0.515$ (see \citealp{Gries2015} for a linguist-friendly introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
The rate of correct predictions is $0.855$, which means a proportional reduction of error of $\lambda=0.354$.
The measure lemma intercepts have a standard deviation of $\sigma_{\text{Measurelemma}}=0.623$, the kind lemma intercept $\sigma_{\text{Kindlemma}}=0.488$.

The coefficient estimates are specified in Table~\ref{tab:bigtable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \textit{confint.merMod} function from \textit{lme4}) with 1,000 replications, and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:bigtable} in the columns labelled \textit{CI Low} and \textit{CI High} (= upper and lower 2.5th percentiles).
The column \textit{CI contains 0} shows an asterisk for those intervals that do \textit{not} include 0.
Furthermore, for each regressor, a $p$ value was obtained by dropping the regressor from the full model, re-estimating the nested model and comparing it to the full model.
Instead of inexact Wald approximations or Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
These bootstrapped $p$ values are given in the columns labelled p\Sub{PB} in Table~\ref{tab:bigtable}.
All regressors have low p\Sub{PB} values of magnitude $10^{-2}$ or $10^{-3}$.
Only \textit{Kindfreq} ($p_{\text{PB}}=0.066$) can be seen as slightly too high to be convincing (`non-significant').

Finally, I show that Bayesian methods in the form of Markov-Chain Monte Carlo estimation do not necessarily (and in this case predictably) lead to different results.
Both the MN model and the F model were re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
The algorithm was run with $4$ chains and $1,000$ iterations, and instead of guessing prior distributions or using implausible uniform prior distributions \cite[251--252]{Levshina2016}, I used plausible default priors.
Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than $10$ or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
\footnote{Consider that with a coefficient of $10$, each increase by $1$ in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
To reliably estimate such coefficients, extremely large samples would be required.}
The algorithm converged, and for all coefficients, the $\hat{R}$ diagnostic was exactly $1$.
The resulting coefficients and intervals as well as an * are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
Both methods lead to exactly the same results (minus negligible numerical differences) as expected (see Section~\ref{sec:rightstatistics}).
The signs and magnitudes of the coefficients are identical, and confidence intervals have the same width and symmetry properties.
If one were interested in significance at $\alpha=0.05$, then the decision based on the confidence intervals differs for \textit{Kindfreq} ($p_{\text{PB}}=0.066$, MLE lower bound is $-0.008$ and MCMC lower bound is $0.009$).
With MCMC, the regressor is `not significant' at the standard $\alpha$ level, but it is with MLE.
This might only go to show how absurd strict significance testing is in complex modelling situations:
both methods essentially estimate the same sign and magnitude for the coefficient (MLE $0.161$, MCMC $0.173$) and do not differ substantially with regard to the precision of the estimate in the form of the width of the confidence interval (both intervals have a width of exactly $0.341$).
Hence, I only interpret the MLE model in the next section.

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/corpus_Measureattraction}~\includegraphics[width=0.5\textwidth]{../R/output/corpus_Kindattraction}
  \caption{Effect plots for the regressors \textit{Measureattraction} and \textit{Kindattraction}; y axes are not aligned}
  \label{fig:eff:attraction}
\end{figure}

The results reported in Section~\ref{sec:corpushierarchicalmodel} generally confirm the hypotheses from Section~\ref{sec:analyses}.
First, the prototypicality effect related to the non-alternating \PGCd\ and \NACb\ can be shown, see the effect plots in Figure~\ref{fig:eff:attraction}.%
\footnote{Effect plots were created using the \textit{effects} package \citep{Fox2003}.
They show the changes in probability for the outcome (y axis) dependent on values of a regressor (x axis), assuming typical values of all other regressors.}
The effect is mostly as expected:
if a lemma appears relatively more often in the \PGCd\ (compared to its frequency in the \NACb), the more often the \PGCa\ is chosen over the \NACa\ with this specific lemma.
The effect for measure nouns is stronger, and it was estimated with higher precision.

@@

I now turn to the predicted effect of plural measure nouns and cardinals as quantifiers.
Plurality has a clear effect in the expected direction (plural favours the \PGCa) ($\beta_{\text{MeasurenumberPl}}=0.814$, $p=0.001$ for \textit{Measurenumber}).
Figure~\ref{fig:eff:leftcontext} shows that the left context of the measure noun influences the choice of the variant ($p=0.001$).
Cardinals have a strong tendency to co-occur with the \NACa\ in both models.
